Social Participation Ontology: community documentation, enhancements and use examples

### 2017-04-25 plosOne/plos.tex
sintonizado abstract

\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
    Participatory democracy advances in virtually all governments and especially in South America which presents a mixed culture and social predisposition.
 In 2012, civil, academic and governmental parties started elaborating the ``Social Participation Common Vocabulary'' (\vcps\ from the Brazilian name \emph{Vocabul\'ario Comum de Participa\c{c}\~ao Social}), as a public and online process. By May 2013, first reference documents were publicized, together with preliminary \owl\ code, logos, and a diagram for a general ``public consultation''.
The \corais\ platform kept online records of the process, like discussions and text drafts. 
This article presents this material and proposes, based on it, the ``Social Participation Ontology'' (\ops\ from the Brazilian name \emph{Ontologia de Participa\c{c}\~ao Social}). To create  this new ontology, these steps were followed: correction of ontological contradictions and \owl\ errors in \vcps; inclusion of concepts documented in the \corais\ platform (but not present in \vcps) in a preliminary version of \ops; entity name standardization; translation of labels to Portuguese, Spanish and English; an ontology expansion; creation of linked data examples and  a \sparql\ endpoint; compilation of use cases from researchers and public managers. Ongoing work involves further adoption of \ops\ by the official Brazilian federal portal for social participation and  NGOs, and linkage to other ontologies for participation. The \ops\ is being used as an upper ontology, and all classes linked further to \foaf\ and \bfo\ as higher upper ontologies.

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLOS ONE authors please skip this step. 
% Author Summary not valid for PLOS ONE submissions.   
%\section*{Author Summary}
%Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut dui eu ante cursus gravida non sed sem. Nullam sapien tellus, commodo id velit id, eleifend volutpat quam. Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium diam, quis maximus dolor faucibus id. Nunc convallis sodales ante, ut ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.
%
\linenumbers

>>>>

\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
    Participatory democracy advances in virtually all governments and especially in South America
    which exhibits a mixed culture and social predisposition.
    This article presents the ``Social Participation Ontology'' 
    (\ops\ from the Brazilian name \emph{Ontologia de Participa\c{c}\~ao Social})
    implemented in compliance with the Web Ontology Language standard (\owl)
    for fostering social participation, specially in virtual platforms.
    The entities and links of \ops\ were defined based on an extensive collaboration of specialists.
    It is shown that \ops\ is instrumental for information retrieval from the contents of the portal,
    both in terms of the actors (at various levels) as well as mechanisms and activities.
    Significantly, \ops\ is linked to other \owl\ ontologies as an upper ontology
    and via \foaf\ and \bfo\ as higher upper ontologies,
    which yields sound organization and access of knowledge and data.
    In order to illustrate the usefulness of \ops, we present results on ontological expansion
    and integration with other ontologies and data.
    Ongoing work involves further adoption of \ops\ by the official
    Brazilian federal portal for social participation and \ngo s,
    and further linkage to other ontologies for social participation.
% Please keep the Author Summary between 150 and 200 words
% Use first person. PLOS ONE authors please skip this step. 
% Author Summary not valid for PLOS ONE submissions.   
%\section*{Author Summary}
%Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut dui eu ante cursus gravida non sed sem. Nullam sapien tellus, commodo id velit id, eleifend volutpat quam. Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium diam, quis maximus dolor faucibus id. Nunc convallis sodales ante, ut ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.
%
\linenumbers
